### 2025-07-17 correlation.qmd
20250708 - track chunk execution times
20250712 - add interpretation
20250717 - Conclusion sections

@@

The covariance is an unstandardized index of the association between two variables.
The formula for (sample) covariance is in @eq-covariance and is the numerator to the formula for [correlation](#sec-correlationExamplesPearson).

@@

@@

Correlation is an index of the association between variables.
The correlation coefficient ($r$) ranges from −1 to +1, and indicates the sign and magnitude of the association.
Although correlation does not imply causation, identifying associations between variables can still be useful because association is a necessary (but insufficient) condition for causality.

>>>>

This chapter provides an overview of correlation analysis.

@@

The covariance is an unstandardized index of the association between two variables.
It represents the average product of two variables' deviations from their respective means.
The formula for (sample) covariance is in @eq-covariance and is the numerator to the formula for [correlation](#sec-correlationExamplesPearson).

@@

The covariance of a variable with itself is the variable's variance:

$$
\begin{aligned}
  \text{Cov}(x, x) &= \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(x_i - \bar{x})\\
  &= \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})^2 \\
  &= \frac{\sum (x_i - \bar{x})^2}{n-1}\\
  &= s^2
\end{aligned}
$$ {#eq-covarianceToVariance}

@@

Correlation is a standardized index of the association between variables.
The correlation coefficient ($r$) ranges from −1 to +1, and indicates the sign and magnitude of the association.
Although correlation does not imply causation, identifying associations between variables can still be useful because association is a necessary (but insufficient) condition for causality.
The correlation coefficient can be heavily impacted by [outliers](#sec-correlationOutliers) and [restricted range](#sec-correlationRestrictedRange).

### 2025-07-17 data-visualization.qmd
20250708 - track chunk execution times
20250717 - Conclusion sections

@@

>>>>

This chapter provides an overview of principles of data visualization and how to implement them using `R`.

@@

This chapter provided an overview of principles of data visualization and how to implement them using `R`.
Key principles of data visualization include: ensuring adequate contrast, highlighting key elements of interest (and using gray for other elements), reducing clutter, and adding helpful titles and annotations.
We provided examples of data visualizations, including [histograms](#sec-histogram), [density plots](#sec-densityPlot), [rug plots](#sec-densityHistogram), [box-and-whisker plots](#sec-boxAndWhiskerPlot), [violin plots](#sec-violinPlot), [ridgeline plots](#sec-ridgelinePlot), [scatterplots](#sec-scatterplot), [line charts](#sec-lineChart), and [bar plots](#sec-barPlot).

### 2025-07-17 download-football-data.qmd
20250717 - Conclusion sections

@@

>>>>

This chapter provides an overview of how to download and process a wide range of football and fantasy football-related data.
We leverage the data in subsequent chapters for data analysis.

@@

# Conclusion {#sec-downloadFootballDataConclusion}

This chapter provided an overview of how to download and process a wide range of football and fantasy football-related data.
Now that we have downloaded and processed a wide range of data, we can use the data in subsequent chapters for data analysis.

### 2025-07-17 draft.qmd
20250708 - track chunk execution times
20250717 - Conclusion sections

@@

>>>>

This chapter provides an overview of the major types of drafts in fantasy football and various draft strategies.

@@

## Conclusion {#sec-draftConclusion}

The two major draft types are [snake drafts](#sec-draftSnake) and [auction drafts](#sec-draftAuction).
[Auction drafts](#sec-draftAuction) take longer than [snake drafts](#sec-draftSnake) but are more fair because every player is available to every manager, so long as the manager is able and willing to bid enough.
There is no one right draft strategy.
In general, it is helpful to draft players with the best value over other players at their position.

### 2025-07-17 fantasy-football.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - Conclusion sections

@@

@@

>>>>

::: {.content-visible unless-format="html:js"}

@Ly2015 provides a video of a brief introduction to American football.

@@

::: {.content-visible unless-format="html:js"}

@NFLFantasy2020 provides a video of a brief introduction to fantasy football.

@@

# Conclusion {#sec-fantasyFootballConclusion}

This chapter provided a brief primer on (American) football and fantasy football.
Football is the most widely watched sport in the United States.
The goal in football is for a team to score more points than their opponent.
In football, points can be scored by scoring a touchdown (6 points) followed by successfully kicking a point-after-touchdown (1 additional point) or successfully completing a two-point conversion (2 additional points), kicking a field goal (3 points), or getting a safety (2 points).

Fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life NFL players.
The participants or "managers" make decisions such as selecting which players to draft, selecting which players start on a weekly basis, identifying players to pick up from the waiver wire, and making trades with other teams.
Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.
In the game of fantasy football, managers accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
A manager's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.

### 2025-07-17 getting-started.qmd
20250717 - Conclusion sections

The book uses the software `R` [@R-base] for statistical analyses (<http://www.r-project.org>).
R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.

@@

>>>>

The book uses the software `R` [@R-base] for statistical analyses (<http://www.r-project.org>).
R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.
This chapter provides an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.

@@

# Conclusion {#sec-gettingStartedConclusion}

This chapter provided an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.
Most of the code throughout the book that performs data management operations leverages the techniques implemented in this chapter.

### 2025-07-17 index.qmd
20250717 - Conclusion sections

This work would not be possible without the important contributions of many people.
This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of R packages, who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, and Adam Peterson, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
In particular, Andrew Tungate made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] that were necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.

>>>>

This work would not be possible without the important contributions of many people.
This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of R packages, who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, and Adam Peterson, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which is useful for scraping fantasy football projections and calculating projections averaged across sources.
Andrew Tungate also made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] to allow computing fantasy points given projected (or actual) player statistics, which was necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.

### 2025-07-17 intro.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - Conclusion sections

@@

>>>>

::: {.content-visible unless-format="html:js"}

@NFL2014 provides a video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football.

@@

## Conclusion {#sec-introConclusion}

In conclusion, fantasy football provides an exciting lens through which to learn statistics, and `R` is a useful (and free!) software for conducting statistical analysis and creating statistical graphics.

### 2025-07-17 multiple-regression.qmd
20250717 - Conclusion sections

@@

On a weekly basis, target share is computed as the number of targets a player receives divided by the team's total number of targets.
It would be nice to compute it this way for seasonal data, as well; however, for calculating seasonal target share, the `nflfastR` [@R-nflfastR] package currently sums up the player's weekly target share.
Thus, seasonal target share (unlike weekly target share)—as it is currently calculated—does not range from 0–1.
For example, if a player has a weekly target share of 15%, that would be a seasonal (cross-week-summed) target share of 2.55 ($0.15 \times 17 \; \text{games} = 2.55$.

@@

The only terms that are significantly associated with fantasy performance among Wide Receivers are weight and target share, both of which showed a positive association with fantasy points.

@@

Target share has a large [effect size](#sec-practicalSignificance).
All of the other [predictors](#sec-correlationalStudy) have small [effect size](#sec-practicalSignificance).

>>>>

This chapter provides an overview of multiple regression.

@@

On a weekly basis, target share is computed as the number of targets a player receives divided by the team's total number of targets.
It would be nice to compute it this way for seasonal data, as well; however, for calculating seasonal target share, the `nflfastR` [@R-nflfastR] package currently sums up the player's weekly target share.
Thus, seasonal target share (unlike weekly target share)—as it is currently calculated—does not range from 0–1.
For example, if a player has a weekly target share of 15%, that would be a seasonal (cross-week-summed) target share of 2.55 ($0.15 \times 17 \; \text{games} = 2.55$).

@@

The only terms that were significantly associated with fantasy performance among Wide Receivers are weight and target share, both of which showed a positive association with fantasy points.

@@

Target share has a large [effect size](#sec-practicalSignificance).
All of the other [predictors](#sec-correlationalStudy) have a small [effect size](#sec-practicalSignificance).

### 2025-07-17 player-evaluation.qmd
20250717 - Conclusion sections

@@

According to the theoretical perspective to sports performance known as the ecological dynamics approach, successful performance in sports involves the coordination of multiple, intertwined skills that are contextually embedded [@DenHartigh2018].
For instance, a Wide Receiver needs to coordinate skills in speed, route running, jumping, and good hands to create space from defenders and to catch poorly thrown passes in a game.
The ecological dynamics approach considers the important interaction of the player, task, and environment.
Thus, to assess players in a way that is likely to be most predictive of their future performance, it is important for the assessment to retain the player–task–environment interaction [@DenHartigh2018].
For example, it would be valuable to use assessments from games or game-like situations in which players must perform tasks that leverage multiple, intertwined skills and that are similar to the tasks you want to predict.

@@

>>>>

This chapter provides an overview of some of the many factors in player evaluation.
Ultimately, player evaluation involves prediction—predicting how well the player will do in future games.

@@

According to the theoretical perspective to sports performance known as the ecological dynamics approach, successful performance in sports involves the coordination of multiple, intertwined skills that are contextually embedded [@DenHartigh2018].
For instance, a Wide Receiver needs to coordinate skills in speed, route running, jumping, and good hands to create space from defenders and to catch poorly thrown passes in a game.
The ecological dynamics approach considers the important interaction of the player, task, and environment.
Thus, to assess players in a way that is likely to be most predictive of their future performance, it is important for the assessment to retain the player–task–environment interaction [@DenHartigh2018].
For example, it would be valuable to use assessments from games or game-like situations in which players must perform tasks that leverage multiple, intertwined skills and that are similar to the tasks you want to predict.
The extent to which the assessments are more reflective of games or game-like situations, we consider the assessment to have better [ecological validity](#sec-ecologicalValidity).

@@

## Conclusion {#sec-playerEvaluationConclusion}

In conclusion, there are many factors to consider when evaluating players.
You will have less direct access to some factors (e.g., cognitive and motivational factors) than others (e.g., historical performance).
When considering how much weight to give to each of various factors, it is important to evaluate how much [predictive validity](#sec-predictiveValidity) each factor has for predicting a player's successful performance, and to weight it accordingly.
In general, samples tend to be stronger than signs for predicting player performance in sports.
For instance, compared to tests of speed, power, and agility at the NFL Combine, collegiate performance is a stronger predictor of performance in the NFL.
That is, previous sports performance is the best predictor of future performance.
When thinking about a player's value, it can be worth thinking their potential upside and downside.
Whether you value a player with higher upside or higher downside may depend on many factors such as, when drafting, whether you are seeking a starter or bench player, and, when in-season, whether you are projected to win by a large margin in your matchpup.
